Combining checkpointing and data compression\\ for high frequency seismic imaging

### 2018-10-04 paper/paper.tex
Change the dataflow figure to single-column
revised conclusion

\section{Conclusions and Future work}
What are the problems?
The performance we developed might be useful in judging whether compression
would be helpful in a particular scenario without actually building a
complete implementation first. We have tried to keep it application
and platform agnostic in order to maximize its applicability. However
there are a few limitations in this model. Firstly, we assumed that
the compression performance we saw for a single checkpoint is
representative of any checkpoint throughout the simulation. This
assumption might have been completely wrong if made about a checkpoint
in the first few time steps since we consistently saw upto 800x
compression in those cases - since they are mostly zeros. In order for
this assumption to be realistic, we use a checkpoint close to the end
of the simulation so that most of the field is perturbed and hence is
close to the worst case in terms of compression results.

In comparing pure compression with pure checkpointing, we assume that every
checkpoint would be compressed and decompressed. However, in scenarios
where the available memory is only slightly less than the required
memory, an implementation that compresses only a subset of the
checkpoints might outperform the expectations of our model.

Our current experimental setup has two  that was realised in ZFP was the lack of parallelization
in the decompression functions. The ZFP manual talks about the
difficulty in implementing a parallel mode decompressor.

Future work:

This work is focussed on checkpointing in memory alone. In a future
work we would like to extend this performance model to include
checkpointing to disk.

The ZFP algorithm that the majority of this paper was based on, only
supports fields with up to 3 dimensions. Since our field is evolving
in time, we would expect some correlations in the time dimensions that
are currently not being exploited in the compression.

As can be seen in figure \ref{fig:varying_memory}, the lack of
parallel decompression has a significant impact on the overall time to
solution. This strategy would become much more beneficial once we
start seeing faster decompression speeds.

In this work, we have ignored the error that is induced by the lossy
compression - assuming that a tolerance of $10^{-4}$ is acceptable for
the application. This assumption needs to be verified in a formal
analysis of how the error in the field propagates through the
simulation, and how it finally affects the cross-correlated gradient. 
The development of a priori and a posteriori error metrics would
greatly help the further use of lossy compression in similar
problems. 
\begin{itemize}
\item Acceptable error bounds
\item A posteriori and a priori metrics
\item Relationship between compression error and field error, propagation of errors in forward simulation
\item Relationship between compression error and gradient error in cross-correlation
\item What else?
\end{itemize}

>>>>

\section{Conclusions and Future work}
We use lossy compression to reduce the computational overhead of checkpointing
in an adjoint computation used in full waveform inversion, a common method in
seismic imaging applications whose memory footprint commonly exceeds the
available memory size in high performance computing systems. We also developed a
performance model that computes whether or not the combination of compression
and checkpointing will outperform pure checkpointing or pure compression in a
variety of scenarios, depending on the available memory size, computational
intensity of the application, and compression ratio and throughput of the
compression algorithm.

Our current result has several limitations that we plan to address in future
work:
\begin{itemize}
\item We do not discuss the accuracy of the results after decompression. This
depends on the application, compression algorithm, and affects the achievable
compression ratios. Our performance model only requires knowledge of the
compression time and ratio, and it is up to the user of this model to determine
what accuracy is needed and thus what compression ratio is realistic for their
application.
\item We only present results using the ZFP compression library. Again, the
performance model is independent of this, but it would be interesting to try a
greater variety of compression algorithms.
\item ZFP currently only supports serial decompression. If ZFP supported
parallel decompression, our experiments would likely show a geater region in
which the combined method is faster than pure checkpointing. Furthermore, ZFP
only supports fields with up to three dimensions, and it is possible that
exploiting similarities between fields at different time steps would yield a
better compression ratio.
\item Our performance model is based on uniform compression ratios and times.
However, many applications, including FWI, are likely to have initial conditions
that contain little information and are easily compressed, and the compression
ratio gradually declines as the field becomes more complex over time. We have
based our experiments on the final wave field, which is presumably the most
difficult to compress.
\item In comparing pure compression with pure checkpointing, we assume that
every checkpoint would be compressed and decompressed. However, in scenarios
where the available memory is only slightly less than the required memory, an
implementation that compresses only a subset of the checkpoints might outperform
the expectations of our model. 
\item We do not discuss multi-level checkpointing, where some checkpoints are
stored on a slower, larger device such as a solid state drive. We expect
compression to be particularly beneficial in these scenarios due to the reduced
data transfer sizes.
\end{itemize}
